---
title: "Week 4 Data Analysis Script"
author: "Sandra"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
```


## Overview

This report presents preliminary data analysis for the Week 3 Examples Data. Specifically, this will report basic descriptive statistics as well as report tests of the following predictions:

1. People will feel more guilt when: 
  + 1.1 They caused the shock (vs. their partner)
  + 1.2 Their partner was morally good (vs. bad)
  + 1.3 The shock was stronger (high > medium > low)
2. Women will report more guilt than men
3. Agreeable people will report more guilt
4. Neurotic people will report more guilt

and

5. People will be more generous when:
  + 5.1 They caused the shock (vs. partner)
  + 5.2 The partner was morally good (vs. bad)
  + 5.3 The shock was stronger (high > medium > low)
6. Women will be more generous than men
7. Agreeable people will be more generous
8. Neurotic people will be more generous



```{r load}
#Load the data
data<-read.csv("Week 3/Processed Data/Week 3 Data PROCESSED.csv")

#Recode Variables:
data$PIN<-as.factor(data$PIN)
data$sex<-as.factor(data$sex)
data$shock<-factor(data$shock,levels=c("low","medium","high"))
data$shockCause<-as.factor(data$shockCause)
data$pMoral<-as.factor(data$pMoral)
```





## Descriptives

Descriptive statistics for guilt:
```{r descGuilt}

c(summary(data$guilt),"SD"=sd(data$guilt,na.rm=T))


```


Descriptive statistics for generosity:
```{r descGenerosity}
#Generosity before the shock:
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm=T))

#Generosity after the shock:
c(summary(data$postShare),"SD"=sd(data$postShare,na.rm=T))

```



Descriptive statistics for personality:
```{rPers}
#Descriptives for agreeableness
c(summary(data$agree),"SD"=sd(data$agree,na.rm=T))

#Descriptives for neuroticism
c(summary(data$neuro),"SD"=sd(data$neuro,na.rm=T))
```




## Guilt Analyses

### Prediction 1.1

Prediction 1.1 was supported. People felt more guilty when they caused the shock than when their partner did:
```{r p1.1}

#Run the t-test for prediction 1.1
p1.1<-t.test(data$guilt~data$shockCause)

#Output the result
p1.1

```


### Prediction 1.2

Prediction 1.2 was supported. People felt more guilty when their partner was morally good:
```{r p1.2}
#Run the t-test for prediction 1.2
p1.2<-t.test(data$guilt~data$pMoral)

#Output the result
p1.2
```





### Prediction 1.3

Prediction 1.3 was supported. People felt more guilty when their partner received a stronger shock:
```{r p1.3}
#RUn the one-way ANOVA for prediction 1.3
p1.3 <-lm(guilt~shock,data=data)
#Output the result
anova(p1.3)

#Output the group means
tapply(data$guilt,data$shock,function(x) mean(x,na.rm=T))
```





### Prediction 2

Prediction 2 was not supported. Women did not report more guilt than men.
```{r p2}
#Run the t-test for prediction 2
p2<-t.test(data$guilt~data$sex)

#Output the result
p2
```



### Prediction 3

Prediction 3 was supported. Agreeable people were more guilty:
```{r p3}

#Runs the correlation test for prediction 3
p3<-cor.test(data$guilt,data$agree)

#Output the results
p3
```



### Prediction 4

Prediction 4 was not supported. Neurotic people were not more guilty:
```{r p4}

#Runs the correlation test for prediction 4
p4<-cor.test(data$guilt,data$neuro)

#Output the results
p4

```



## Generosity Analyses

```{r melt}
#Melt our data from wide format to long format
dataMelt<-melt(data,
               measure.vars=c("preShare", "postShare"),
               variable.name=c("prePost"),
               value.name=c("share")
               )

```





### Prediction 5.1

Prediction 5.1 was supported. People were more generous when they caused the shock and *less* generous when partner caused shock:
```{r p5.1}
#Run within-subjects ANOVA for p5.1
p5.1<-aov(share~prePost*shockCause+Error(PIN),data=dataMelt)

#Output the results:
summary(p5.1)

#Output the means:
tapply(dataMelt$share,dataMelt[,c(24,21)],function(x) mean(x,na.rm=T)
       )
```

